Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-05 paper/tv.tex
no compilation errors
m

We investigate an alternative approach for modeling and verifying 
compositional real-time systems. In this work, 
we propose a novel temporal specification language, 
which enables a compositional verification via a  Hoare-style 
forward verifier and a term rewriting system (TRS). 
More specifically, we specify system behaviours in the form of 
{Timed Effects} (\timedEffects), which integrates the Kleene Algebra with dependent values and arithmetic constraints, 
%gaining the expressive power beyond finite-state machines; and (ii) introduces a new parallel operator 
to 
%capture the parallelism between event-triggered (sync) and  time-triggered (async) execution traces, 
provide real-time abstractions into traditional linear temporal logics. 
%Resource usages, provided by a \emph{dynamic ticks}\cite{von2017real}, incorporating the time-triggered (physical time) and event-triggered (logical time) execution loops. 
For example, one safety property, \textit{"The event \anyevent{Done} 
will be triggered no later than one second"}\footnote{For 
simplicity, we use integer values to represent seconds in this 
paper, while it can bes extended to real numbers and other 
time measurement units.}, is expressed in \timedEffects\ as: 
\begin{align*}
&\code{\effect \defeq \  0 {\leq} t {<}1 : ({\_^\star} \cdot \anyevent{Done} ) \# t}.
\end{align*}  
%The timed effects incorporates different kinds of existing temporal logics, such as linear-time temporal logic (LTL)
%and .
%which replaces temporal operators by time-constrained versions from 
Here, \code{\#} is a novel operator specifying the \emph{real-time} 
constraints for the \emph{logical-time} sequences \cite{von2017real}; 
\code{\_} is a wildcard matching to any event; 
Kleene star \code{\star} denotes trace repetition.
The above formula \code{\effect} corresponds to `\code{\Diamond_{[0, 1)}\ }\anyevent{Done}' 
in metric temporal logic (MTL), reads \textit{"within one second, 
\anyevent{Done} finally happens"}. Moreover, the time bounds can be dependent on the program inputs. For example, we express, the effects of \code{send(d)} as:
\begin{align*}
&\code{\effect^{send (d)} \defeq \  (0 {<} d {\leq}5 \wedge  0 {\leq} t {<} d) : ({\anyevent{Send} \# t)} \cdot \anyevent{Done}}.
\end{align*}  
The \code{send} method takes a parameter \code{d}, and \anyevent{Send}s 
out a message at time \code{t}.  
The above formula  \code{\effect^{send (d)}} indicates that the 
input  \code{d} is positive and smaller or equal to \code{5}; the method 
generates a finite trace, with the first event \anyevent{Send}, followed by 
a second event \anyevent{Done}; and \anyevent{Send} takes a no more 
than \code{d} milliseconds delay to finish. 
Although these examples are simple, they show the benefits of deploying 
value-dependent 
time bounds. Intuitively, if traditional timed automata define an 
\emph{exact} transition system, \timedEffects\ define 
a set (possibly infinite) of exact transition systems.

   %they already illustrate the gain of expressiveness from  existing temporal logics and traditional timed automata.

%properties beyond

@@

Existing automatic verification techniques massively rely on 
the timed automata which requires manually cast of clock variables with 
carefully calculated clock constraints.

@@

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   ((x:=v)^*, {meth^*})} 
%\qquad   \qquad\quad 
 %   \m{(Basic\ Types)} \qquad  \code{b} \code{\quad  ::= \quad }
  %   \code{int\ |\ bool\ |\ void}
    \\

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{ mn  \ {(x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Values)} & \code{v} &\code{  ::= } & ()\ |\ c \ | \ b \ |\ x\  
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ v \ | \ 
    [b]\ e \ |\ mn({v^*}) \ }
  \code{| 
 % v.f:=e \ | 
 \ e_1;e_2\ | \ e_1 || e_2 \ 
  |\ {if}\ b\ {then}\ e_1\ {else}\ e_2 \
 }  \\
&&& \code{ |\ \textbf{event} [\anyevent{A}(v)] \{(x{:=}v)^*\} }
 \\
 &   & & \code{
  | \ \textbf{delay} [d] \ 
  | \ e_1\ \textbf{timeout} [d] \  e_2 \ 
  | \ \textbf{deadline}  [d] \ e \  
  }  \\
  ~\\

@@

\\
\m{(Events)} & ev  & \ ::= \ & 
\code{
  \anyevent{A}(c, \pi, (x:=v)^*)
    \ | \ \anynotevent{A} 
    \ | \ \_  
}

%    \\
%        \m{(Real\ Time)} &  \mathcal{O} &\code{::= }&
%    
%    \code{
%   %n \ | \ 
%   c \ | \ (c, {+}\infty) \ | \  [0, c) \ | \  \mathcal{O}_1 \wedge \mathcal{O}_2 \ | \  \mathcal{O}_1 \vee \mathcal{O}_2
%    }

@@

%The core implementation is less than 3500 LOC, roughly, 1{+} LOC for the verifier, and 2{+} LOC for the TRS. We prove the soundness of the TRS. 
%Next, we show case studies to demonstrate the feasibilities of our \timedEffects comparing to timed automata.

@@

Our implementation is the first that solves the  inclusion problem of symbolic timed automata, which is considered important because it overcomes the following main limitations in traditional timed model checking:  i) they cannot be used to specify and verify systems incompletely specified (i. e., whose timing constants are not known yet), and hence cannot be used in early design phases; ii) verifying a system for a set of timing constants usually requires to enumerate all of them one by one if they are supposed to be integer-valued; in addition, timed automata cannot be used to verify a system for a set of timing constants that are to be taken in a real-valued dense interval.

@@

%
%
%compare to timed automata:
%modular and can plug in anywhere. 
%absolute time and relative time. 
%$
%//
%(lower_t, upper_time)
%lower_ti<=upper_t
%lower_t<5
%lower_t>8
%modular and can plug in anywhere. 
%absolote time and relative time. 
%it can be only for synchronous system, t.sync, 
%or or real time t.rt.dur. 
%$

>>>>

We investigate an alternative approach for modeling and verifying 
compositional real-time systems. In this work, 
we propose a novel temporal specification language, 
which enables a compositional verification via a  Hoare-style 
forward verifier and a term rewriting system (TRS). 
More specifically, we specify system behaviours in the form of 
{Timed Effects} (\timedEffects), which integrates the Kleene Algebra with dependent values and arithmetic constraints, 
%gaining the expressive power beyond finite-state machines; and (ii) introduces a new parallel operator 
to 
%capture the parallelism between event-triggered (sync) and  time-triggered (async) execution traces, 
provide real-time abstractions into traditional linear temporal logics. 
%Resource usages, provided by a \emph{dynamic ticks}\cite{von2017real}, incorporating the time-triggered (physical time) and event-triggered (logical time) execution loops. 
For example, one safety property, \textit{"The event \anyevent{Done} 
will be triggered no later than one time unit"}\footnote{Without loss of generality, 
we use integer values to represent time units in this 
paper, while it can bes extended to real numbers and other 
time measurement units.}, is expressed in \timedEffects\ as: 
\begin{align*}
&\code{\effect \defeq \  0 {\leq} t {<}1, ({\_^\star} \cdot \anyevent{Done} ) \# t}.
\end{align*}  
%The timed effects incorporates different kinds of existing temporal logics, such as linear-time temporal logic (LTL)
%and .
%which replaces temporal operators by time-constrained versions from 
Here, \code{\#} is a novel operator specifying the \emph{real-time} 
constraints for the \emph{logical-time} sequences \cite{von2017real}; 
\code{\_} is a wildcard matching to any event; 
Kleene star \code{\star} denotes trace repetition.
The above formula \code{\effect} corresponds to `\code{\Diamond_{[0, 1)}\ }\anyevent{Done}' 
in metric temporal logic (MTL), reads \textit{"within one second, 
\anyevent{Done} finally happens"}. Moreover, the time bounds 
can be dependent on the program inputs. For example, as the program 
shown in \autoref{fig:Value_dependent_intro}, we mark the precondition and postcondition
using keywords \code{\emph{require}} and \code{\emph{ensure}} respectively.

\newcommand{\codem}[1]{{\code{\emph{#1}}}}

\begin{wrapfigure}{L}{0.4\columnwidth}
  \vspace{-2mm}
\begin{lstlisting}
void addSugar (int n) 
/*require n>=0 , _^*
  ensure  t>=n, Sugar#t */
{ if (n == 0) { 
    event ["Sugar"];} 
  else {
    timeout (oneSugar() , 1);
    addSugar (n-1);}} 
\end{lstlisting}
\vspace{-1mm}
\caption{Value-dependent specification.} 
\label{fig:Value_dependent_intro}
\vspace{-1mm}
\end{wrapfigure}

The function 
\code{\emph{addSugar}} is intended to takes a parameter \codem{n}, representing 
the portion of the sugar we need to add. When \codem{n} decreases to \codem{0}, 
it simply raises an event \anyevent{Sugar} to mark the end of the process. 
Otherwise, it adds one portion of the sugar then recursively calls \code{\emph{addSugar}} with parameter \codem{n-1}. 
Statement \codem{timeout(e, d)} sets a timer which executes code \codem{e} once the timer expires at time \codem{d}. 
Therefore, the time spent on adding one portion of the sugar is no less than 1 time unit.

The specifications of  \codem{addSugar(n)} indicates that the 
input  \code{n} is non-negative; the method 
generates a finite trace where \anyevent{Sugar} takes a no less 
than \code{d} milliseconds delay to finish. 
Although these examples are simple, they show the benefits of deploying 
value-dependent 
time bounds, which is beyond the capability of Timed Automata. Intuitively, if traditional Timed Automata define an 
\emph{exact} transition system, \timedEffects\ define 
a set (possibly infinite) of exact transition systems.

\begin{comment}
  we express, the effects of \code{send(d)} as:
\begin{align*}
&\code{\effect^{send (d)} \defeq \  (0 {<} d {\leq}5 \wedge  0 {\leq} t {<} d),  ({\anyevent{Send} \# t)} \cdot \anyevent{Done}}.
\end{align*}  
effects of \code{addSugar(n)} as:
\code{(n{>}0  \wedge  t {\geq} n) : ({\anyevent{Sugar} \# t)}}

\end{comment}

   %they already illustrate the gain of expressiveness from  existing temporal logics and traditional Timed Automata.

%properties

@@

Existing automatic verification techniques massively rely on 
the Timed Automata which requires manually cast of clock variables with 
carefully calculated clock constraints.

@@

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   (\alpha^*, {meth^*})} 
%\qquad   \qquad\quad 
 %   \m{(Basic\ Types)} \qquad  \code{b} \code{\quad  ::= \quad }
  %   \code{int\ |\ bool\ |\ void}
    \\
    \m{(Assignment)} &  \code{\alpha }&\code{  ::= }&
    \code{x:=v} 
   \\

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{ mn  \ {(x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Values)} & \code{v} &\code{  ::= } & ()\ |\ c \ | \ b \ |\ x\  
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ v \ | \ 
    [b]\ e \ |\ mn({v^*}) \ }
  \code{| 
 % v.f:=e \ | 
 \ e_1;e_2\ | \ e_1 || e_2 \ 
 % |\ {if}\ b\ {then}\ e_1\ {else}\ e_2 \
 |\ \textbf{event} [\anyevent{A}(v, \alpha^*)] 
 }  \\

 &   & & \code{
  | \ \textbf{delay} [d] \ 
  | \ e_1\ \textbf{timeout} [d] \  e_2 \ 
  | \ \textbf{deadline}  [d] \ e \  
  }  \\
  ~\\

@@

\\
\m{(Events)} & ev  & \ ::= \ & 
\code{
  \anyevent{A}(v, \pi, \alpha^*)
    \ | \ \anynotevent{A} 
    \ | \ \_  
}

@@

%The core implementation is less than 3500 LOC, roughly, 1{+} LOC for the verifier, and 2{+} LOC for the TRS. We prove the soundness of the TRS. 
%Next, we show case studies to demonstrate the feasibilities of our \timedEffects comparing to Timed Automata.

@@

Our implementation is the first that solves the  inclusion problem of symbolic Timed Automata, which is considered important because it overcomes the following main limitations in traditional timed model checking:  i) they cannot be used to specify and verify systems incompletely specified (i. e., whose timing constants are not known yet), and hence cannot be used in early design phases; ii) verifying a system for a set of timing constants usually requires to enumerate all of them one by one if they are supposed to be integer-valued; in addition, Timed Automata cannot be used to verify a system for a set of timing constants that are to be taken in a real-valued dense interval.

@@

%
%
%compare to Timed Automata:
%modular and can plug in anywhere. 
%absolute time and relative time. 
%$
%//
%(lower_t, upper_time)
%lower_ti<=upper_t
%lower_t<5
%lower_t>8
%modular and can plug in anywhere. 
%absolote time and relative time. 
%it can be only for synchronous system, t.sync, 
%or or real time t.rt.dur. 
%$
